---
title: "Statistical Analyses"
author: "PJKA"
date: "19/06/2020"
output: github_document
editor_options: 
  chunk_output_type: console
---
#Setup

##Install Packages and Import Data

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Load required libraries
library(tidyverse)
library(vegan)
library(ggfortify)
library(rcartocolor)
library(patchwork)
library(EcolUtils)


#Import original dataset from .csv
EnvDOM_v2<- read.csv("Aukes et al LO Letters Dataset.csv")

#Average values so only one value per site:
avg_all2 <- EnvDOM_v2 %>%
  select("Env", "Location","Ecozone","Area","Type","LocType","Sample", "DOC","Hphi","BP","HS","BB","LMWN","LMWA","SUVA") %>%
  group_by(Env, Location,Ecozone,Area,Type,LocType,Sample) %>%
  summarise_all(mean) %>%
  ungroup()
```

##Conform Database  

Based on previous analyses, we are removing high-organic subsurface samples as they skew the analysis (i.e.: we know they are different so lets remove them for now and see how all the other samples compare).  

```{r, echo=FALSE, warning=FALSE, message=FALSE}
pca_avg_all2 <- avg_all2 %>%
                  filter(Type != 'Wet. Subsurface')

#Create a dataframe to recalculate LC-OCD output so eluted components are proportions of overall eluted DOM (mgC/L) concentration
#select parameters:
pca_avg_all2 <- pca_avg_all2 %>%
  transmute(Env = Env,
            Location = Location,
            Ecozone = Ecozone,
            Area = Area,
            Type = Type,
            LocType = LocType,
            Sample = Sample,
            DOC = DOC,
            BP = BP/Hphi,
            HS = HS/Hphi,
            BB = BB/Hphi,
            LMWN = LMWN/Hphi,
            LMWA = LMWA/Hphi)

```

##Statistical Analyses
###Create PCA
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Create a matrix of LC-OCD proportions for input into PCA 
lcocd.matrix <-pca_avg_all2[,9:13]

#run PCA
lcocd.pca<- prcomp(lcocd.matrix, center=TRUE, scale.=TRUE)

#look at output of PCA
summary(lcocd.pca)
screeplot(lcocd.pca, type="lines")

#create vectors to associate 'rows' with actual sample information
lcocd.pca.Env<-pca_avg_all2[,1]
lcocd.pca.Location<-pca_avg_all2[,2]
lcocd.pca.Ecozone<-pca_avg_all2[,3]
lcocd.pca.Type<-pca_avg_all2[,5]
lcocd.pca.LocType<-pca_avg_all2[,6]
lcocd.pca.Sample<-pca_avg_all2[,7]

#Associated statistics for PCA
#To see loadings:
lcocd.pca$rotation

#To see individual sample scores:
lcocd.pca$x

#To get eigenvalues:
lcocd.pca$sdev^2

#this package offers a nice way to obtain correlation between variables and PC
library(FactoMineR)
lcocd.pca2 <- PCA(lcocd.matrix)

#eigenvalue:
lcocd.pca2$eig

#corr b/w variables and PC
lcocd.pca2$var$coord
```

###Plot PCA
```{r, warning=FALSE, echo=FALSE, message=FALSE}
autoplot(lcocd.pca, scale = 0, 
         data=pca_avg_all2, 
         fill = 'Type', shape= 'Ecozone', size =2,
         loadings=TRUE, loadings.colour = 'blue',
         loadings.label= TRUE, loadings.label.colour = 'black',loadings.label.size=3)+
  scale_fill_carto_d(palette = 'Vivid')+
  scale_shape_manual(values=c(21,24,23,22,25))+
  guides(fill= guide_legend(override.aes = list(shape=21)) )+
  theme_bw()+
  theme(panel.grid = element_blank())
```

###PERMANOVA Calculation

Here we ask whether hydrologic compartments are significantly different:

```{r}
adonis2(lcocd.matrix ~ Type, data = pca_avg_all2, permutations = 999, method='euclidean', strata='Ecozone')
```

As the p-value<0.01, we reject the null hypothesis and can say that there is a significant difference in DOM composition between hydrologic compartments.

###Post hoc comparisons

The PERMANOVA indicates that there are statistical differences, but does not show where. Here we conduct a pairwise a postereri test to show significance between means of different hydrologic compartments using a Dunn's Test.

```{r}
#Using EcolUtils package
adonis.pair(vegdist(pca_avg_all2[,9:13]), as.factor(pca_avg_all2$Type), corr.method = 'bonferroni', nper = 8000)
```
