---
title: "Statistical Analyses"
author: "PJKA"
output: github_document
editor_options: 
  chunk_output_type: console
---
# Setup

## Install Packages and Import Data

```{r, message=FALSE, warning=FALSE}
#Load required libraries
library(tidyverse)
library(vegan)
library(ggfortify)
library(rcartocolor)
library(patchwork)
library(EcolUtils)


#Import dataset file from WLU Dataverse
EnvDOM <- read.csv("Aukes-DOM_Comp_Data-dataset.csv")

#Average samples to result in one value per sampling site
avg_all <- EnvDOM %>%
  select("Env", "Location","Ecozone","Type","LocType","Sample", "DOC","Hphi","BP","HS","BB","LMWN","LMWA","SUVA") %>%
  group_by(Env, Location,Ecozone,Type,LocType,Sample) %>%
  summarise_all(mean) %>%
  ungroup()

#Rename 'Groundwater' to 'Groundwater'
avg_all <- avg_all %>%
  mutate(Type = replace(Type, Type=='Subsurface', 'Groundwater'),
         Type = replace(Type, Type=='Wet. Subsurface', 'Wet. Groundwater'))

```

## Conform Database  

Based on previous analyses, we are removing high-organic subsurface samples as they skew the analysis (i.e.: we know they are different so lets remove them for now and see how all the other samples compare).  

```{r, message=FALSE, warning=FALSE}
pca_avg_all <- avg_all %>%
                  filter(Type != 'Wet. Groundwater')

#Create a dataframe to recalculate LC-OCD output so eluted components are proportions of overall eluted DOM (mgC/L) concentration
#select parameters:
pca_avg_all <- pca_avg_all %>%
  transmute(Env = Env,
            Location = Location,
            Ecozone = Ecozone,
            Type = Type,
            LocType = LocType,
            Sample = Sample,
            DOC = DOC,
            BP = BP/Hphi,
            HS = HS/Hphi,
            BB = BB/Hphi,
            LMWN = LMWN/Hphi,
            LMWA = LMWA/Hphi)

```

## Statistical Analyses
### Create PCA
```{r, warning=FALSE}
#Create a matrix of LC-OCD proportions for input into PCA 
lcocd.matrix <-pca_avg_all[,8:12]

#run PCA
lcocd.pca<- prcomp(lcocd.matrix, center=TRUE, scale.=TRUE)

#look at output of PCA
summary(lcocd.pca)
screeplot(lcocd.pca, type="lines")

#create vectors to associate 'rows' with actual sample information
lcocd.pca.Env<-pca_avg_all[,1]
lcocd.pca.Location<-pca_avg_all[,2]
lcocd.pca.Ecozone<-pca_avg_all[,3]
lcocd.pca.Type<-pca_avg_all[,4]
lcocd.pca.LocType<-pca_avg_all[,5]
lcocd.pca.Sample<-pca_avg_all[,6]

#Associated statistics for PCA
#To see loadings:
lcocd.pca$rotation

#To see individual sample scores:
lcocd.pca$x

#To get eigenvalues:
lcocd.pca$sdev^2

#this package offers a nice way to obtain correlation between variables and PC
library(FactoMineR)
lcocd.pca2 <- PCA(lcocd.matrix)

#eigenvalue:
lcocd.pca2$eig

#corr b/w variables and PC
lcocd.pca2$var$coord
```

### Plot PCA
```{r, message=FALSE, warning=FALSE}
autoplot(lcocd.pca, scale = 0, 
         data=pca_avg_all, 
         fill = 'Type', shape= 'Ecozone', size =2,
         loadings=TRUE, loadings.colour = 'blue',
         loadings.label= TRUE, loadings.label.colour = 'black',loadings.label.size=3)+
  scale_fill_carto_d(palette = 'Vivid')+
  scale_shape_manual(values=c(21,24,23,22,25))+
  guides(fill= guide_legend(override.aes = list(shape=21)) )+
  theme_bw()+
  theme(panel.grid = element_blank())
```

### PERMANOVA Calculation

Here we ask whether water body type is significantly different. We do not have a fully balanced design to assess the interaction of 'Ecozone', but instead nest it within the function: 

```{r, warning=FALSE}
adonis2(lcocd.matrix ~ Type, data = pca_avg_all, permutations = 999, method='euclidean', strata='Ecozone')
```

As the p-value<0.01, we reject the null hypothesis and can say that there is a significant difference in DOM composition between water body types.

### Post hoc comparisons

The PERMANOVA indicates that there are statistical differences, but does not show where. Here we conduct a pairwise a postereri test to show significance between means of different water body types using a Dunn's Test.

```{r, warning=FALSE}
#Using EcolUtils package
adonis.pair(vegdist(pca_avg_all[,8:12]), as.factor(pca_avg_all$Type), corr.method = 'bonferroni', nper = 8000)
```
